---
title: "Fast Fashion's Rapid Rate of Overproduction is Observedly Unsustainable & Unethical: Why and How did we get here?"
author: "Jayden Jung"
thanks: "Code and data are available at: https://github.com/jj-andj/fast-fashion-analysis"
date: "April 21, 2023"
date-format: long
format: 
  pdf: 
    documentclass: article
    geometry: margin = 1in
    abstract: "Fast fashion is extremely profitable, accessible, and convienient, and equally -- if not more -- unethical and unsustainable. To illustrates the severity of fast fashion companies' rate of overproduction, this paper scrapes a representative company's online inventory at multiple points in time and finds that old products are being retired and new products added at unbelievably fast rates: too fast to be ethically possible. Secondary research explores this industry's cultural origins and catalysts and its serious negative ethical and environment consequences. Possible solutions are discussed, and the paper concludes on calling for urgent awareness and action to combat fast fashion."
output: 
  bookdown::pdf_document2
toc: false
fig_caption: yes
nocite: '@*'
bibliography: references.bib
---

# Introduction

> "Fast fashion isn't free. Someone, somewhere is paying."
>
> --- Lucy Siegle, Journalist and Producer of *The True Cost*, 2015.

Fast fashion is the design, production, and selling of clothing in a very rapid and high volume approach. Companies participating in this practice churn out countless new, trendy products faster than ever at unbelievably cheap prices and even lower costs, reaping profit margins in the hundreds of millions and even billions. Customers are happy, finding it an accessible and affordable way to dress well and constantly update their closets. This situation may appear too good to be true: this is because it is.

As Lucy Siegle's quote suggests, fast fashion does come at a cost, though it may not be obvious. It is made possible only by riding on the backs of unethical labor conditions and catastrophically negative effects on the environment. Despite these truths, many continue to support and buy from fast fashion companies, often as they do not grasp the true severity of these issues or that they feel it is not something that can be challenged.

This paper delves into a case study of the fast fashion company ZAFUL. By documenting all available products in one of ZAFUL's given categories through web scraping and comparing the data collected on different dates over time, we see first-hand the unbelievably fast rate of production and turnover. The extremity of these results are further emphasized when compared to the product retention rate of a sustainable fashion company, KOTN.

The paper then draws on secondary literature to convey how the illustrated severity of fast fashion companies' rapidness in production is inherently unethical and unsustainable, specifically by showcasing the negative consequences that are necessary to achieve those results. Consideration of the cultural origins and catalysts of the fast fashion industry are also provided. Then, discussion of possible solution spaces to these problems are explored, generally suggesting that fast fashion itself should be fought against.

Fast fashion and its ramifications are not as widely understood as it must be for us to societally move away from it. It is this paper's intention to shine light on this topic and call for urgent awareness and action.

The gathering of the data used for this paper was done by web scraping through the statistical programming language `R` [@citeR], the packages `rvest` [@citeVest], `httr` [@citeHTTR], and the Chrome extension *Web Scraper* [@citeWS] which were conducted on the sites zaful.com and kotn.com. Gathering, cleaning, and further analysis was also supported by packages `tidyverse` [@citeTidy], `dplyr` [@citeDplyr], `knitr` [@citeKnit], `kableExtra` [@citeKable], and `RColorBrewer` [@citeRColor].

# Data

talk about data collection process and category (women's clothing only)

web scraper chrome extension for kotn (dynamic webpage)

for looping through pagination for zaful

only property we are using is product name

differences/sameness were found by the exact product name so even the difference in colour or size mean they are diff products (this is a lil unfortunate)

summarize total number of products seen before and after for two stores (table woahh)

```{r}
#| label: set-up
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(RColorBrewer)

#Read in data
z_products <- read_csv(here::here("inputs/data/zafulapril2.csv"), show_col_types = FALSE)
z_products <- z_products[!duplicated(z_products$products),]

z_products2 <- read_csv(here::here("inputs/data/zafulapril4.csv"), show_col_types = FALSE)
z_products2 <- z_products2[!duplicated(z_products2$products),]

z_products3 <- read_csv(here::here("inputs/data/zafulapril10.csv"), show_col_types = FALSE)
z_products3 <- z_products3[!duplicated(z_products3$products),]

k_products <- read_csv(here::here("inputs/data/kotnapril2.csv"), show_col_types = FALSE)
k_products <- k_products[!duplicated(k_products$products),]

k_products2 <- read_csv(here::here("inputs/data/kotnapril4.csv"), show_col_types = FALSE)
k_products2 <- k_products2[!duplicated(k_products2$products),]

k_products3 <- read_csv(here::here("inputs/data/kotnapril10.csv"), show_col_types = FALSE)
k_products3 <- k_products3[!duplicated(k_products3$products),]

```

```{r}
#| label: tbl-data
#| tbl-cap: "Total number of Women's Clothing products by store by date"
#| echo: false
#| message: false
#| warning: false

# create data frame of all total products by date by store
totals <- data.frame(
  store = c("ZAFUL (Fast Fashion)", "KOTN (Sustainable)"),
  apr2 = c(nrow(z_products), nrow(k_products)),
  apr4 = c(nrow(z_products2), nrow(k_products2)),
  apr10 = c(nrow(z_products3), nrow(k_products3))
)

# create table displaying data by the dates & store
totals %>% 
  knitr::kable(booktabs = T, linesep = "\\addlinespace", col.names = c("Store", "April 2", "April 4", "April 10")) %>% 
  column_spec(1, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position")%>% 
  kable_styling(latex_options = "striped")

```

```{r}
#| label: fig-data
#| fig-cap: "ZAFUL holds much more products overall than KOTN"
#| echo: false
#| message: false
#| warning: false

# create data frame of all information by date and store
totals2 <- data.frame(
  store = c(rep("ZAFUL (Fast Fashion)", times = 3), 
            rep("KOTN (Sustainable)", times = 3)),
  date = c(rep(c("April 2", "April 4", "April 10"), times = 2)),
  products = c(nrow(z_products), nrow(z_products2), nrow(z_products3),
               nrow(k_products), nrow(k_products2), nrow(k_products3))
)

# rename columns and values
date = c("April 2", "April 4", "April 10")
totals2$date <- factor(totals2$date, levels = date)

# create graph of data 
totals2 %>% 
  ggplot(aes(x = date, y = products, fill = store)) +
  geom_col(position = "dodge") + 
  theme_minimal() +
  labs(x = "Date", y = "Products", fill = "Store") + 
  theme(text=element_text(size=11, family="serif")) + 
  scale_fill_brewer(palette = "Set1")

```

# Results

```{r}
#| label: tbl-summary
#| tbl-cap: "Breakdown of added and deprecated products by date shows severity of ZAFUL's rate of production"
#| echo: false
#| message: false
#| warning: false

# Creating subsets of what products are same, added, and removed 
## KOTN Apr 2 -> Apr 4
k_same <- subset(k_products2, (products %in% k_products$products))
k_added <- subset(k_products2, !(products %in% k_products$products))
k_removed <- subset(k_products, !(products %in% k_products2$products))

## KOTN Apr 2 -> Apr 10
k_same2 <- subset(k_products3, (products %in% k_products$products))
k_added2 <- subset(k_products3, !(products %in% k_products$products))
k_added2_1 <- subset(k_products3, !(products %in% k_products2$products))
k_removed2 <- subset(k_products, !(products %in% k_products3$products))
k_removed2_1 <- subset(k_products2, !(products %in% k_products3$products))

## ZAFUL Apr 2 -> Apr 4
z_same <- subset(z_products2, (products %in% z_products$products))
z_added <- subset(z_products2, !(products %in% z_products$products))
z_removed <- subset(z_products, !(products %in% z_products2$products))

## ZAFUL Apr 2 -> Apr 10
z_same2 <- subset(z_products3, (products %in% z_products$products))
z_added2 <- subset(z_products3, !(products %in% z_products$products))
z_added2_1 <- subset(z_products3, !(products %in% z_products2$products))
z_removed2 <- subset(z_products, !(products %in% z_products3$products))
z_removed2_1 <- subset(z_products2, !(products %in% z_products3$products))

#create columns for data frame 
stores <- c("ZAFUL", "", "", "", "", "KOTN", "", "", "", "")
statuses <- c("Total", 
             "Added from April 2", 
             "Deprecated from April 2", 
             "Added from April 4", 
             "Deprecated from April 4", 
             "Total", 
             "Added from April 2", 
             "Deprecated from April 2", 
             "Added from April 4", 
             "Deprecated from April 4")
april2 <- c(nrow(z_products),
           '',
           '',
           '',
           '',
           nrow(k_products),
           '',
           '',
           '',
           '')
april4 <- c(nrow(z_products2),
           nrow(z_added),
           nrow(z_removed),
           '',
           '',
           nrow(k_products2),
           nrow(k_added),
           nrow(k_removed),
           '',
           '')
april10 <- c(nrow(z_products3),
           nrow(z_added2),
           nrow(z_removed2),
           nrow(z_added2_1),
           nrow(z_removed2_1),
           nrow(k_products3),
           nrow(k_added2),
           nrow(k_removed2),
           nrow(k_added2_1),
           nrow(k_removed2_1))

# create data frame of all added and removed data
summary <- data.frame(
  store = stores,
  status = statuses,
  apr2 = april2,
  apr4 = april4,
  apr10 = april10)

# create kable of data frame
summary %>% 
   knitr::kable(booktabs = T, linesep = "\\addlinespace", col.names = c("Store", " ", "April 2", "April 4", "April 10")) %>% 
  column_spec(1:2, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  kable_styling(latex_options = "striped")
```

```{r}
#| label: tbl-percent
#| tbl-cap: "ZAFUL's percentage of additions and deprecations of products"
#| echo: false
#| message: false
#| warning: false

inc1 <- ((nrow(z_added))/(nrow(z_products)) * 100)
inc2 <- ((nrow(z_added2_1))/(nrow(z_products2)) * 100)
inc3 <- ((nrow(z_added2))/(nrow(z_products)) * 100)

dec1 <- ((nrow(z_removed))/(nrow(z_products)) * 100)
dec2 <- ((nrow(z_removed2_1))/(nrow(z_products2)) * 100)
dec3 <- ((nrow(z_removed2))/(nrow(z_products)) * 100)

percent <- data.frame(
  status = c("Increase (+%)", "Decrease (-%)"),
  apr2to4 = c(inc1, dec1),
  apr2to10 = c(inc2, dec2),
  apr4to10 = c(inc3, dec3)
)

percent %>% 
  knitr::kable(booktabs = T, linesep = "\\addlinespace", col.names = c(" ", "April 2nd to 4th", "April 4th to 10th", "April 2nd to 10th")) %>% 
  column_spec(1, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position")%>% 
  kable_styling(latex_options = "striped")
```

```{r}
#| label: fig-added
#| fig-cap: "Total num of products seem consistent, but the proportion of added ones are large"
#| echo: false
#| message: false
#| warning: false

dates <- c("April 2", "April 4", "April 10")
productss <- c(nrow(z_products), 
               (nrow(z_products2) - nrow(z_added)), 
               (nrow(z_products3) - nrow(z_added2)), as.integer(0), 
               nrow(z_added), 
               nrow(z_added2))

added <- data.frame(
  date = c(rep(dates, times = 2)),
  add = c(rep("Same from Apr 2nd", times = 3), rep("Added", times = 3)),
  products = productss
)

date = c("April 2", "April 4", "April 10")
added$date <- factor(totals2$date, levels = date)

added %>% 
  ggplot(aes(x = date, y = products, fill = add)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Date", y = "Products", fill = "Added?") + 
  theme(text=element_text(size=11, family="serif")) + 
  scale_fill_brewer(palette = "Set1")
```

# Discussion

## Findings

Lorem

## Why did we get here? Exploring culture-based roots of Fast Fashion

Lorem

## At what cost? Reviewing the ramifications of Fast Fashion

Lorem

## Defining Solution Spaces

Lorem

## Future Research

Lorem

## Limitations

Lorem

\clearpage

# References {.unnumbered}
